---
title: "Day 3 Homework"
output: html_document
---

The goal of this homework is to get practice with manipulating data.

```{r}
GSS_sub <- read.csv("C:/Users/HYH/Desktop/WINTER 2015/R_programming/GSS_subset7212.csv")
myGSS <- GSS_sub
library(reshape2)
summary(myGSS)
head(myGSS)
```

1. Make both a "wide" and a "long" version of your data

```{r}
## wider data.frame
myGSS <- GSS_sub
myGSS$RACE <- as.factor(myGSS$RACE)
levels(myGSS$RACE) <- c("White","Black","Other")
myGSS.cast <- dcast(myGSS, CASEID+FAIR+TRUST+YEAR+AGE+SEX+HISPANIC+EDUC+DEGREE+MARITAL ~ RACE, value.var = "HELPFUL")
colnames(myGSS.cast)[11] <- "White_Helpful"
colnames(myGSS.cast)[12] <- "Black_Helpful"
colnames(myGSS.cast)[13] <- "Other_Helpful"
head(myGSS.cast)

## longer data.frame
myGSS <- GSS_sub
head(myGSS)
myGSS.melt <- melt(myGSS, id.vars=c("CASEID","YEAR","AGE","SEX","RACE","HISPANIC","EDUC","DEGREE","MARITAL"), value.name="ordinalscale", variable.name="measurement.type")
head(myGSS.melt)
tail(myGSS.melt)
```


## -casting with a LOt of variables
```{r eval=FALSE}

mydata.wide <- dcast(mydata.long, [80 columns here] ~ split)

columns.to.keep <- colnames(mydata.long)[1:80]
columns.to.keep.pasted <- paste(colnames.to.keep, collapse = "+")
my.cast.formula <- as.formula(paste(columns.to.keep.pasted, "~split"))
class(my.cast.formula) # "formula"
mydata.wide <- dcast(mydata.long, formula = my.cast.formula)

```

2. Transform a variable

```{r}
myGSS <- GSS_sub
myGSS$AGE2 <- myGSS$AGE/10
head(myGSS)
summary(myGSS)
```

3. Recode a factor

```{r}
myGSS <- GSS_sub
myGSS$MARITAL[myGSS$MARITAL==9] <- NA
summary(myGSS)
myGSS$MARITAL <- as.factor(myGSS$MARITAL)
levels(myGSS$MARITAL) <- c("Married","Widowed","Divorced","Separated","Never Married")
head(myGSS)
summary(myGSS)
```

4. Add a new factor, where the levels are conditional on the values of some other variable. E.g., a factor of "high", "medium", and "low" based on thresholds for another variable.

```{r}
myGSS <- GSS_sub
myGSS$AGE[myGSS$AGE>89 ] <- NA
myGSS$AGECat<-cut(myGSS$AGE, br=c(0,31,61,90),  labels=c(1:3))
levels(myGSS$AGECat) <- c("under 31", "31-60", "61 up")
head(myGSS)
summary(myGSS)
```

5. Reorder the levels of this new factor

```{r}
myGSS$AGECat <- relevel(myGSS$AGECat, "61 up")
levels(myGSS$AGECat)
head(myGSS)
```

6. Choose a factor that could be re-coded as multiple dimensions (recall the sepal/petal vs. length/width breakdown). If you don't have one, try to create one.  Now create two (or more) new factors that represent these dimensions.

```{r}
myGSS <- GSS_sub
head(myGSS)
myGSS.1 <- subset(myGSS, select=c("CASEID", "SEX", "RACE"))
head(myGSS.1)

myGSS.1$YOUNG.ARM <- (rnorm(57061)^2)*20
myGSS.1$OLD.ARM <- (rnorm(57061)^2)*21
myGSS.1$YOUNG.LEG <- (rnorm(57061)^2)*23
myGSS.1$OLD.LEG <- (rnorm(57061)^2)*22

myGSS.2 <- melt(myGSS.1, id.vars=c("CASEID","SEX", "RACE"), value.name="cm", variable.name="measurement.type")
head(myGSS.2)

myGSS.3 <- cbind(myGSS.2, colsplit(myGSS.2$measurement.type, pattern="\\.", names=c("Seniority", "BodyPart")))
head(myGSS.3)
tail(myGSS.3)
```

7. Using `dcast`, aggregate some variable into "cell means" of some kind.

```{r}
myGSS.agg1 <- dcast(myGSS.3, CASEID+SEX+RACE ~ BodyPart, value.var="cm", fun.aggregate = mean)
head(myGSS.agg1)
tail(myGSS.agg1)

myGSS.agg2 <- dcast(myGSS.3, SEX+RACE ~ BodyPart, value.var="cm", fun.aggregate=mean)
head(myGSS.agg2)
tail(myGSS.agg2)
```

8.  Save all of the above in a well-documented R Markdown document.  Email me the .Rmd file (and if you use different or updated data from Day 2 Homework, send that to me, too).


## try `ifelse` & `if` function