---
title: "Day 4 Practice"
output: html_document
---

1. Read in the data base of phoneme pairs from Wedel, Kaplan, and Jackson (2013).  The relative path from the working directory of this project is: "scottdata/CognitionPaperModifiedData.csv"

```{r eval=TRUE}
mydata <- read.csv("C://Users/HYH/Desktop/projects/r_programming_bsos/scottdata/CognitionPaperModifiedData.csv")
summary(mydata)
```

2. Read in the data using the WRONG function.  What does it look like?

mydata <- read.table("scottdata/CognitionPaperModifiedData.csv")
summary(mydata)
mydata <- read.delim("scottdata/CognitionPaperModifiedData.csv")
summary(mydata)

3. Change some (or all) of the column names.  Provide a table that links original column names with the current/new column names.

```{r}
oldnames <- colnames(mydata)
mycolnames <- colnames(mydata)
mycolnames[7:17] <- c("Minpairs.ALL", "Minpairs.BETWEEN", "Minpairs.WITHIN", "Entropy.ALL", "Entropy.WITHIN", "Entropy.SEG", "Entropy.BIPHONE","SegTok.HI","SegTok.LO","SegType.HI","SegType.LO")
data.frame(oldnames, mycolnames)
```

4. Locate NAs.  Remove all rows containing NAs.

```{r}
summary(mydata)
mydata.na <- apply(mydata, 1, function(x){any(is.na(x))})
summary(mydata.na)
mydata.rmna <- mydata[!mydata.na,]
summary(mydata.rmna)

# mydata.rmna <- mydata[complete.cases(mydata),]
```

5. Recode the "Corpus" variable in the following way:
Am = American
Du = Dutch
Fr = French
Ge = German
HK = Cantonese
Ko = Korean
RP = British
Sl = Slovakian
Sp = Spanish

```{r}
summary(mydata)
levels(mydata$Corpus)
levels(mydata$Corpus) <- c("American", "Dutch", "French", "German", "Cantonese", "Korean", "British", "Slovakian", "Spanish")
levels(mydata$Corpus)
head(mydata)
```

6. Run a logistic regression, using `glm()`, where "Merged" status is the (binary) dependent variable, and "WithinMinPairs" is the predictor.
  - Hint #1: the basic usage of `glm()` is:
```{r eval=FALSE}
results <- glm(DV ~ predictor1 + predictor2 + ...  , 
               data = dataframe.object,
               family = "binomial")
```
  - Hint #2: you need the DV to be 0s and 1s, not a factor.

```{r eval=TRUE}
summary(mydata)
class(mydata$Merged)
mydata$DV <- ifelse(mydata$Merged == "Merged", 1, 0)  ## create a new numeric variable
# mydata$DV2 <- as.numeric(mydata$Merged) - 1 ## wrong order, but a possibility.  a bit risky
class(mydata$DV)
mydata.logit <- glm(DV~WithinMinPairs, data=mydata, family="binomial")
summary(mydata.logit)
```

7. Use `summary()` to display the results.

summary(mydata.logit)

8. Using INLINE code, report the estimate, standard error, and p-value for the WithinMinPairs coefficient.  This will look something like:

```{r}
mydata.sum <- summary(mydata.logit)
names(mydata.sum)
mydata.sum$coefficients
class(mydata.sum$coefficients)
mydata.sum$coefficients[2,1]
Est. <- mydata.sum$coefficients["WithinMinPairs", "Estimate"]
# Std.err <- mydata.sum$coefficients["WithinMinPairs", "Std.Error"]  what does subscript-out-of-bounds mean?
Std.err <- mydata.sum$coefficients[2,2]
p.value <- mydata.sum$coefficients["WithinMinPairs", "Pr(>|z|)"]
data.frame(Est., Std.err, p.value)

Est. <- round(Est., digits=4)
Std.err <- round(Std.err, digits=4)
p.value <- round(p.value, digits=4)
data.frame(Est., Std.err, p.value)
```

# The estimate for the WithinMinPairs coefficient was `r round(est, 4)`, with a standard error of `r std.err ` and a p of `r round(p-value, 3)`.

EXTRA CREDIT: figure out how to display the above results with only a few decimal places

HINT: use `str()` to figure out how to extract these values.

