---
title: "Day 6 Practice, Loops"
output: pdf_document
---

1. Create a chunk of code that does the following:
  - get a subset of data from the `iris` data set that just includes one species
  - get the mean of each of the values, and arrange these in a one-row data frame
  - make a PDF plot of Sepal.Width by Petal.Width

```{r}
head(iris)
summary(iris)
myiris <- subset(iris, c(Species=="setosa"))
head(myiris)

mean.values <- c(mean(myiris[[1]], na.rm=TRUE), mean(myiris[[2]], na.rm=TRUE), mean(myiris[[3]], na.rm=TRUE), mean(myiris[[4]], na.rm=TRUE))
mydf <- data.frame(x=colnames(myiris[1:4]), y=mean.values)
mydf

library(ggplot2)
pdf("day6_practice_loops_plot_HYH.pdf", width=8, height=5)
ggplot(myiris, aes(x=Sepal.Width, y=Petal.Width))+geom_point(color="darkblue", size=3)+theme_minimal()
dev.off()
```

2. Now make a copy of this code and turn it into a loop:
  - loop through all three species
  - combine the results from each loop into a single data frame
    - do this the "slow" way and the "fast" way
  - combine the plots into a single PDF

```{r}
myiris <- iris
myiris$Group <- rep(1:3, each=50)
summary(myiris)

thisvector <- myiris[myiris$Group==1, "Sepal.Length"]
myanswer<- mean(thisvector)

for(thisgroup in 1:3) {
  thisvector.1 <- myiris[myiris$Group==thisgroup, "Sepal.Length"]
  myanswer.1 <- mean(thisvector.1)
  thisvector.2 <- myiris[myiris$Group==thisgroup, "Sepal.Width"]
  myanswer.2 <- mean(thisvector.2)
  thisvector.3 <- myiris[myiris$Group==thisgroup, "Petal.Length"]
  myanswer.3 <- mean(thisvector.3)
  thisvector.4 <- myiris[myiris$Group==thisgroup, "Petal.Width"]
  myanswer.4 <- mean(thisvector.4)
  print(c(myanswer.1,myanswer.2,myanswer.3,myanswer.4))}

rm(myanswer)
myanswer <- 1
for(thisgroup in c(1,2,3)) {
  thisvector <- myiris[myiris$Group==thisgroup, "c(Sepal.Length, Sepal.Width)"]
  myanswer[thisgroup] <- mean(thisvector)}

rm(myanswer)
myanswer <- vector(length=4)
for(thiscounter in 1:3) {
  thisgroup <- c(1,2,3,4)[thiscounter]
  thisvector <- myiris[myiris$Group==thisgroup, "Sepal.Length"]
  myanswer[thiscounter] <- mean(thisvector)
  print(myanswer)[thiscounter]}


```

3. Make another copy of the loop, adding an `if` statement that will have the effect of skipping plants where `Petal.Width` is less than 1.