---
title: "Day 7 Homework"
output: html_document
---

1. Work on making progress on your Final Project.

GSS_sub <- read.csv("C:/Users/HYH/Desktop/WINTER 2015/R_programming/GSS_subset7212.csv")
gss.sub <- GSS_sub
summary(gss.sub)

# Recode missing & top-coded values to NAs
```{r}
gss.sub$HELPFUL[gss.sub$HELPFUL==0 | gss.sub$HELPFUL==8 | gss.sub$HELPFUL==9] <- NA
gss.sub$FAIR[gss.sub$FAIR==0 | gss.sub$FAIR==8 | gss.sub$FAIR==9] <- NA
gss.sub$TRUST[gss.sub$TRUST==0 | gss.sub$TRUST==8 | gss.sub$TRUST==9] <- NA
gss.sub$AGE[gss.sub$AGE >89 ] <- NA
gss.sub$EDUC[gss.sub$EDUC >20 ] <- NA
gss.sub$DEGREE <- gss.sub$DEGREE+1
gss.sub$DEGREE[gss.sub$DEGREE > 5] <- NA
gss.sub$MARITAL[gss.sub$MARITAL==9] <- NA
summary(gss.sub)
```

# Remove rows containing NAs
```{r}
summary(gss.sub) # 50761 cases in total
gss.sub[!complete.cases(gss.sub),]
mygss <- na.omit(gss.sub)
summary(mygss)
nrow(mygss) # 34379 cases left
```

# Create age groups
```{r}
mygss$AGECAT <- cut(mygss$AGE, br=c(0,31,61,90),  labels=c(1:3))
levels(mygss$AGECAT) <- c("18-31", "31-60", "61 up")
```

# Create social trust scale
```{r}
install.packages("car")
library(car)
mygss$HELPFUL <- recode(mygss$HELPFUL, '"1"="3"; "2"="1"; "3"="2"')
mygss$FAIR <- recode(mygss$FAIR, '"2"="3"; "3"="2"')
mygss$TRUST <- recode(mygss$TRUST, '"1"="3"; "2"="1"; "3"="2"')

mygss$SOCIAL.TRUST <- (mygss$HELPFUL + mygss$FAIR + mygss$TRUST) / 3
summary(mygss)
```

2. Document some road blocks you are running into.

3. Email them to me before 7am tomorrow.