---
title: "Day 9 Homework"
output: html_document
---

```{r}
load("C:/Users/HYH/Desktop/projects/r_programming_bsos/cleaned-up data.RData")
summary(mygss)
library(reshape2)
library(dplyr)

```

1. Write a function to do a bunch of clean-up and formatting to your data.

```{r}
cor_pval <- function(v1, v2) {
  cor_test <- cor.test(v1, v2, alternative=c("two.sided"), method=c("pearson"))
  cor <- cor_test$estimate
  pval <- cor_test$p.value
  if (pval < .05 & cor > 0) { output <- paste("r is =", round(cor, 4), "p < .05, positive correlations!") } else if (pval < .05 & cor < 0) { output <- paste("r is =", round(cor, 4), "p < .05, negative correlations!") }
  else { output <- paste("r is =", round(cor, 4), "p is =", round(pval, 4), "no significant correlations") }
  output }

cor_pval(mygss$AGE, mygss$SOCIAL.TRUST)

```

2. Make sure this function works on your data.

head(mygss)
cor_pval(mygss$EDUC, mygss$SOCIAL.TRUST)

3. Get some subset of your data, or find a similar-but-not-identical data set.  Change up the column order.  Re-run your function on this new data.  Does it work?

```{r}
mygss2 <- mygss %>%
  select(YEAR, SEX, HELPFUL, EDUC, MARITAL, AGE, SOCIAL.TRUST, RACE) %>%
  filter(SEX==2, RACE!=3, YEAR>=1986, AGE>30)
summary(mygss2)

cor_pval(mygss2$AGE, mygss2$SOCIAL.TRUST)
# Yes, it works. 

```

4. Fix your function to be able to work on the new data.

5. Write a function to do some kind of summary stat, and use the `dplyr` functions to run this function over groups of your data.

```{r}
educ.boost <- function (x) { median(x)^2 }

results <- mygss %>%
  group_by(SEX, AGECAT) %>%
  summarize(EDUCboost = educ.boost(EDUC))
results

```
